Brief notes on erosion

### 2022-05-06 docs/PhysicsPaper/StokesErosion.tex
NM

% APS journals: PRL and PRF
% Note: All AIP and APS journals use the revtex package; both PRL and PRF are APS.
\documentclass[reprint, superscriptaddress, notitlepage]{revtex4-1}

% PACKAGES
\usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
\usepackage{comment, enumerate, tabularx, multirow}
\usepackage{natbib, hyperref, url}
\usepackage{algorithm, algpseudocode, pifont, longtable, subfigure}
%\usepackage[justification=RaggedRight]{caption}

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% COMMANDS
% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
\newcommand{\HERE}[1]{ \vsp{#1} {\color{blue} HERE} }
\newcommand{\tofix}{\color{blue}}
\newcommand{\new}{}		% \color{green}
\newcommand{\newa}{}		% \color{red}
\newcommand{\newb}{}		% \color{blue}
% Basic math, derivatives
\newcommand{\td}[2]{\frac{d #1 }{d #2}}
\newcommand{\ttd}[2]{\frac{d^2 #1 }{{d #2}^2}}
\newcommand{\pd}[2]{ \frac{ \partial #1}{ \partial #2 } }
\newcommand{\ppd}[2]{ \frac{ \partial^2 #1}{ {\partial #2}^2 } }
% Basic math, vectors and other
\newcommand{\BigO}[1]{ \mathcal{O} \left(#1\right) }
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\grad}{\nabla} 
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\tavg}[1]{\langle #1 \rangle}
\newcommand{\norm}[1]{ \left\| #1 \right\| }
\newcommand{\ip}[1]{ \langle #1 \rangle }
% For real and imaginary, could use \Re or \Im, or \mathcal{R}, or \text{Re}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

\author{M.~N.~J.~Moore}
\affiliation{Department of Mathematics and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}
\author{Bryan D.~Quaife}
\affiliation{Department of Scientific Computing and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}
\author{Shang-Huan Chiu}
\affiliation{Department of Scientific Computing, Florida State
University, Tallahassee, FL, 32306.}

@@

@@

%Maybe cite: \cite{Rycroft2016}

@@

\bibliographystyle{plain}
\bibliography{refs}
\end{document}

>>>>

% APS journals: PRL and PRF
% Note: All AIP and APS journals use the revtex package; both PRL and PRF are APS.
\documentclass[preprint, superscriptaddress, notitlepage]{revtex4-1}

%\documentclass{jfm}
% NOTE: jfm documentclass needed for \upi
%---------------------------------------------------%
% Packages from JFM 2020
%\usepackage[top=1.2in,bottom=1.2in,left=1in, right=1in]{geometry}
\usepackage{amsfonts, amssymb, array}
\usepackage[fleqn,reqno]{amsmath}
\usepackage{graphics, graphicx, subfigure}
% Replaced subcaption with subfigure
\usepackage{todonotes, comment, soul, float}
%---------------------------------------------------%

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% COMMANDS
% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
% For real and imaginary, could use \Re or \Im, or \mathcal{R}, or \text{Re}
\newcommand{\Real}{\Re}
\newcommand{\Imag}{\Im}

%---------------------------------------------------%
% From JFM 2020
\newcommand{\bd}{{\partial}}
\newcommand{\bigO}{{\mathcal{O}}}
\newcommand{\cc}{{\mathbf{c}}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\DD}{{\mathcal{D}}}
\newcommand{\DDD}{{\boldsymbol{\mathcal D}}}
\newcommand{\eeta}{{\boldsymbol\eta}}
\newcommand{\ff}{{\mathbf{f}}}
\newcommand{\grad}{{\nabla}}
\newcommand{\II}{{\mathbf{I}}}
\newcommand{\iin}{\mathrm{in}}
\newcommand{\llambda}{{\boldsymbol\lambda}}
\newcommand{\nn}{{\mathbf{n}}}
\newcommand{\NN}{{\mathcal{N}}}
\newcommand{\out}{\mathrm{out}}
\newcommand{\rr}{{\mathbf{r}}}
\renewcommand{\Re}{{\operatorname{Re}}}
\renewcommand{\Im}{{\operatorname{Im}}}
\newcommand{\RR}{{\mathbb{R}}}
\renewcommand{\ss}{{\mathbf{s}}}
\newcommand{\ssigma}{{\boldsymbol\sigma}}
\newcommand{\tar}{\mathrm{tar}}
\newcommand{\bary}{\mathrm{bary}}
\newcommand{\trap}{\mathrm{trap}}
\newcommand{\uu}{{\mathbf{u}}}
\newcommand{\UU}{{\mathbf{U}}}
\newcommand{\vv}{{\mathbf{v}}}
\newcommand{\xx}{{\mathbf{x}}}
\newcommand{\xxi}{{\boldsymbol{\xi}}}
\newcommand{\yy}{{\mathbf{y}}}
\newcommand{\mcaption}[2]{\caption{\small \em #1}\label{#2}} 
\newcommand{\secref}[1]{\ref{#1}}
\def\gap{\hspace*{.2in}}
% Nick's below
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ppd}[2]{\frac{\partial^2 #1}{{\partial #2}^2}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\Vn}{V_\nn}
\newcommand{\Vs}{V_\ss}
\newcommand{\CE}{C_E}
%---------------------------------------------------%

@@

\author{Nicholas J.~Moore}
\affiliation{Department of Mathematics and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}
\author{Bryan D.~Quaife}
\affiliation{Department of Scientific Computing and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}
\author{Shang-Huan Chiu}
\affiliation{Department of Scientific Computing, Florida State University, Tallahassee, FL, 32306.}

@@

%Maybe cite: \cite{Rycroft2016}
% Cite paper by Ladd "Wormhole formation in dissolving fractures"

@@

@@

% NOTE: THIS SECTION HAS BEEN PLAGIARIZED FROM JFM 2020
% SO WE NEED TO PARAPHRASE IT
%\begin{comment}
%-----------------------------------------------------------------------------------------------%
\section{Governing Equations}
\label{sec:formulation}
%-----------------------------------------------------------------------------------------------%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\citep{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of the boundary of $[-3,3] \times [-1,1]$.  All eroding
bodies are placed in $[-1,1] \times [-1,1]$ to create a buffer region
that allows the flow profile imposed at the inlet to transition to the
more complex flow intervening between the bodies. Neglecting inertial
forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, and $\Vn$ is the normal
velocity of $\gamma$. Since the rate of erosion is much slower than the
background flow, a quasi-steady approximation justifies the no-slip
boundary condition. Parameters include the fluid viscosity $\mu$ and the
material-dependent erosion constant $\CE$. By non-dimensionalizing
equations~\eqref{eqn:erosionModel}~\citep{qua-moo2018}, we set both
these parameters to one. The shear stress on $\gamma$ is
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
where, $\nn$ is the normal vector pointing into the body, and $\ss$ is
the unit tangent vector pointing in the counterclockwise direction. We
simulate erosion by alternating between solving the fluid equations and
advancing the eroding grains.  The strength of $\UU$ is adjusted at each
time step to achieve a constant pressure drop across the channel,
motivated by the geological situation of a porous medium connecting two
regions of  fixed hydraulic heads.

%-----------------------------------------------------------------------------------------------%
\section{Boundary Integral Equation Formulation}
\label{sec:DLP}
%-----------------------------------------------------------------------------------------------%
To accurately solve the governing equations~\eqref{eqn:erosionModel} in
complex two-dimensional geometries, we reformulate the equations as a
BIE.  This has the advantage that only the one-dimensional boundary of
the domain must be discretized, and, with appropriate quadrature
formulas and fast summation methods, the result is a high-fidelity
numerical simulation with near-optimal computational complexity.

%-----------------------------------------------------------------------------------------------%
\subsection{Double-Layer Potential Formulation in $\RR^2$}
%-----------------------------------------------------------------------------------------------%
Applying the same approach as our previous work~\citep{qua-moo2018}, we
start with the double-layer potential 
\begin{align}
  \DDD[\eeta](\xx) = \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:velocityDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  We complete the integral equation
formulation by adding the $M$ Stokeslets,
$S[\llambda_\ell,\cc_\ell](\xx)$, and $M$ rotlets,
$R[\xi_\ell,\cc_\ell](\xx)$, where $\cc_\ell$ is a point inside the
$\ell^{th}$ body~\citep{pow-mir1987}.  Here $\llambda_\ell$ and
$\xi_\ell$ are the Stokeslet and rotlet strengths, respectively,
corresponding to the $\ell^{th}$ body.  Then, for any sufficiently
smooth geometry $\Omega$, the solution of the incompressible Stokes
equation with a Dirichlet boundary condition $\ff$ is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:BIE}
\begin{alignat}{3}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \NN_0[\eeta](\xx) \nonumber \\
    &\quad + \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), 
    \quad &&\qquad\xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy, &&\qquad \ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy, 
    &&\qquad \ell = 1,\ldots,M.
\end{alignat}
\end{subequations}
Here, the null space associated with the flux-free condition of $\ff$ is
addressed with  $\NN_0$ which is the integral operator with kernel
$N_0(\xx,\yy) = \nn(\xx) \otimes \nn(\yy)$, $\xx,\yy \in \Gamma$.  In
this work, $\ff$ is the prescribed velocity, which is equal to $\UU$ on
the outer wall, $\Gamma$, and equal to zero on the eroding bodies,
$\gamma_\ell$, $\ell=1,\ldots,M$.

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
corresponding deformation tensor, pressure, and vorticity at $\xx \in
\Omega$ are written in terms of layer potentials~\citep{qua-moo2018}.
To compute the deformation tensor for $\xx \in \gamma$, we include the
jump term
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right].
  \label{eqn:deformationJump}
\end{align}
Finally, the deformation tensor, pressure, and vorticity due to the
Stokeslets and rotlets are readily available~\citep{poz1992}. Having
computed the deformation tensor on $\gamma$, the shear stress is
computed using equation~\eqref{eqn:shearStress}.

%-----------------------------------------------------------------------------------------------%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%-----------------------------------------------------------------------------------------------%
The velocity double-layer potential~\eqref{eqn:velocityDLP}, and its
corresponding deformation tensor, pressure, and vorticity are all
written as layer potentials in $\RR^2$.  However, the quadrature method
we introduce in section~\ref{sec:method} requires complex-valued
representations. The first step to form a complex representation is to
write the Laplace double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Real (v(x)),
\end{align}
where
\begin{align}
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega.
  \label{eqn:laplaceComplex}
\end{align}
Here $x = x_1 + i x_2,y = y_1 + i y_2 \in \CC$ are the complex
counterparts of $\xx = (x_1,x_2),\yy = (y_1,y_2) \in \RR^2$, and $\eta =
\eta_1 + i \eta_2$ is the complex counterpart of $\eeta =
(\eta_1,\eta_2)$. Therefore, depending on the formulation of the layer
potential, $\Omega$ is interpreted as a subset of $\RR^2$ or $\CC$.
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy integral by
first finding the boundary data of $v$. If $\Omega$ is a
simply-connected interior domain, then the boundary data of $v$
satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \bd\Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Having computed the boundary data of the
holomorphic function $v$, by the Cauchy integral theorem we have
\begin{subequations}
  \label{eqn:cauchy}
  \begin{alignat}{3}
  v(x) &= \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \\
  v'(x) &= \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \\
  v''(x) &= \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
  \end{alignat}
\end{subequations}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

Finally, the Stokes double-layer potential~\eqref{eqn:velocityDLP} can
be written using a Laplace double-layer
potential~\eqref{eqn:laplaceComplex} and its gradients 
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Therefore, the Stokes double-layer potential can be written as a sum of
Cauchy integrals and its first derivative~\citep{bar-wu-vee2015}
\begin{equation}
  \begin{aligned}
    u_1(x) &= \Real (v[\psi_1](x)) + \Real (v'[y\cdot\eta](x)) 
             -x_1\Real (v'[\eta_1](x)) - x_2\Real (v'[\eta_2](x)), \\
    u_2(x) &= \Real (v[\psi_2](x)) - \Imag (v'[y\cdot\eta](x)) 
         +x_1\Imag (v'[\eta_1](x)) + x_2\Imag (v'[\eta_2](x)),
  \end{aligned}
  \label{eqn:cauchyVelocity}
\end{equation}
where $y \cdot \eta = y_1 \eta_1 + y_2 \eta_2$, 
\begin{align} 
  \psi_1=(\eta_1+i\eta_2)\frac{\Real(n)}{n}, \quad
  \psi_2=(\eta_1+i\eta_2)\frac{\Imag(n)}{n},
\end{align}
and $n \in \CC$ is the complex counterpart of the outward unit normal
$\nn \in \RR^2$.

%-----------------------------------------------------------------------------------------------%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%-----------------------------------------------------------------------------------------------%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Real (v'[\psi_1](x)) + 
    \Real (v''[y\cdot\eta](x)) - \Real (v'[\eta_1](x)) \\
    &- x_1\Real (v''[\eta_1](x)) - x_2\Real (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Imag (v'[\psi_1](x)) - 
    \Imag (v''[y\cdot\eta](x)) + x_1\Imag (v''[\eta_1](x)) \\
    &- \Real (v'[\eta_2](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Real (v'[\psi_2](x)) - 
    \Imag (v''[y\cdot\eta](x)) + \Imag (v'[\eta_1](x))  \\
    &+ x_1\Imag (v''[\eta_1](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Imag (v'[\psi_2](x)) - 
    \Real (v''[y\cdot\eta](x)) + x_1\Real (v''[\eta_1](x)) \\
    &+ \Imag (v'[\eta_2](x)) + x_2\Real (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \Omega$, the Cauchy integral representation of the
vorticity at $x \in \Omega$ is
\begin{align}
  \omega(x) = 
    \Real (v'[\psi_2](x)) + \Imag (v'[\psi_1](x))+ 
    \Real (v'[\eta_2](x))+ \Imag (v'[\eta_1](x)).
\end{align}
%\end{comment}

%\bibliographystyle{plain}
%\bibliography{refs}
\end{document}

### 2022-05-06 docs/TexNotes/Notes.tex
NM

\subsection{Isolating the resistivity due to the bodies}
The above total resistivity reflects the resistivity from the porous matrix as well as the channel walls. There is a buffer region that lies outside the porous-matrix region ($x_0$ lies in the buffer region), and we would like to remove the contribution from this region. Lets regard the total resistivity as an area-integral of resistivity-density, so that we can write the total in terms of the areas of the regions
\begin{equation}
A_{t} r_{t} = A_{b} r_{b} + A_{m} r_{m}
\end{equation}
Here, the subscripts are as follows: $t$ stands for total, $b$ for buffer, and $m$ for matrix. Then we can solve for the resistivity in the matrix region
\begin{equation}
r_m = \frac{A_t r_t - A_b r_b}{A_m} = x_0(r_t - 3) + 3
\end{equation}
In the buffer region, resistivity is only due to channel walls so that $r_b = 3$. The area of the included buffer region depends on $x_0$ via $A_b = 4(x_0 - 1)$. The area of the matrix region is always $A_m = 4$, and $A_t = A_b + A_m = 4 x_0$. The resulting resistivity in the matrix, $r_m$, includes the contribution from the bodies as well as the channel walls. We can then subtract the contribution from the walls to get $\tilde{r}_m$, the resistivity due to the bodies alone as
\begin{equation}
\tilde{r}_m = r_m - 3 \, .
\end{equation}
Putting everything together, the final formula is
\begin{equation}
\label{resmatrix0}
\tilde{r}_m = x_0 (r_t - 3) \, .
\end{equation}
As convoluted as this calculation seems, it does appear to give convergent numerical results that are {\em independent of the choice of $x_0$} (to the expected order of accuracy).

@@

\subsection{Relationship between resistivity and total drag}

Consider the stress tensor
\begin{equation}
\bvec{\sigma} = -p \bvec{I} + \mu \left( \grad \bu + \grad \bu^T \right)
\end{equation}
Below we will set $\mu=1$. The Stokes equation is
\begin{equation}
\grad \cdot \bvec{\sigma} = 0
\end{equation}
By the divergence theorem we have
\begin{equation}
\label{divthm}
0 = \int_{\Omega} \grad \cdot \bvec{\sigma} \, dV 
= \int_{\partial \Omega} \bvec{\sigma} \cdot \nhat \, dS
\end{equation}
for any domain $\Omega$. Now consider $\Omega$ to be a section of the channel $-x_0 < x < x_0$ and spanning the entire width $-w < y < w$. Then the integral on the right-hand-side of \eqref{divthm} has two contributions: first, the total drag arising from all immersed bodies as well as from the top and bottom walls, and second, the difference between the upstream and downstream pressures (we have implicitly assumed that $x_0$ is sufficiently distant from the bodies so that $\grad \bvec{u}$ is nearly zero). Since the total integral above vanishes, we have the approximate relationship
\begin{equation}
F_D \approx 2 w \left( \pup - \pdn \right)
\end{equation}
where $F_D$ is the total drag force from all bodies and from the top and bottom walls. The error in this approximation is due to neglecting the contribution of $\grad \bvec{u}$ in the stress tensor at the upstream and downstream positions. We thus have for the permeability
\begin{equation}
k_{tot} = \frac{2 x_0 \, \qavg}{\pup-\pdn} \approx 4 x_0 w \qavg F_D^{-1}
\end{equation}
and the resistivity
\begin{equation}
r_{tot} = \frac{\pup-\pdn}{2 x_0 \, \qavg} \approx \frac{F_D}{4 x_0 w \qavg} = \frac{3 F_D}{8 x_0 w U}
\end{equation}
Where we have used $\qavg = \frac{2}{3} U$. These are the permeability and resistivity for the entire region $-x_0 < x < x_0$. In particular, the resistivity of the porous matrix will be averaged with the much lower resistivity of the buffer region. To remove the contribution from the buffer region, and thus isolate the resistivity due to the porous matrix of bodies, we can simply set $x_0 = 1$ above. With also setting $w=1$, we get
\begin{equation}
\label{resdrag}
r_{tot} = \frac{3}{8} \frac{F_D}{U}
\end{equation}
With this simplification, I see a match between measuring resistivity via the drag \eqref{resdrag} versus measuring resistivity directly via \eqref{resmatrix}.

>>>>

%-------------------------------------------------------------%
% ISOLATING RESISTIVITY DUE TO BODIES
%-------------------------------------------------------------%
\subsection{Isolating the resistivity due to the bodies}
The above total resistivity reflects the resistivity from the porous matrix as well as the channel walls. There is a buffer region that lies outside the porous-matrix region ($x_0$ lies in the buffer region), and we would like to remove the contribution from this region. Let's regard the total resistivity as an area-integral of resistivity-density, so that we can write the total in terms of the areas of the regions
\begin{equation}
A_{t} r_{t} = A_{b} r_{b} + A_{m} r_{m}
\end{equation}
Here, the subscripts are as follows: $t$ stands for total, $b$ for buffer, and $m$ for matrix. Then we can solve for the resistivity in the matrix region
\begin{equation}
r_m = \frac{A_t r_t - A_b r_b}{A_m} = x_0(r_t - 3) + 3
\end{equation}
In the buffer region, resistivity is only due to channel walls so that $r_b = 3$. The area of the included buffer region depends on $x_0$ via $A_b = 4(x_0 - 1)$. The area of the matrix region is always $A_m = 4$, and $A_t = A_b + A_m = 4 x_0$. The resulting resistivity in the matrix, $r_m$, includes the contribution from the bodies as well as the channel walls. We can then subtract the contribution from the walls to get $\tilde{r}_m$, the resistivity due to the bodies alone as
\begin{equation}
\tilde{r}_m = r_m - 3 \, .
\end{equation}
Putting everything together, the final formula is
\begin{equation}
\label{resmatrix0}
\tilde{r}_m = x_0 (r_t - 3) \, .
\end{equation}
As convoluted as this calculation seems, it does appear to give convergent numerical results that are {\em independent of the choice of $x_0$} (to the expected order of accuracy).

@@

% New May 5, 2022
\subsubsection{Modification for free-slip BCs}
For the free-slip BCs, the resistivity of the buffer region vanishes, $r_b = 0$, giving
\begin{equation}
r_m = \frac{A_t}{A_m} r_t = x_0 r_t
\end{equation}

%-------------------------------------------------------------%
% RESISTIVITY AND DRAG
%-------------------------------------------------------------%
\subsection{Relationship between resistivity and total drag}

Consider the stress tensor
\begin{equation}
\bvec{\sigma} = -p \bvec{I} + \mu \left( \grad \bu + \grad \bu^T \right)
\end{equation}
Below we will set $\mu=1$. The Stokes equation is
\begin{equation}
\grad \cdot \bvec{\sigma} = 0
\end{equation}
By the divergence theorem we have
\begin{equation}
\label{divthm}
0 = \int_{\Omega} \grad \cdot \bvec{\sigma} \, dV 
= \int_{\partial \Omega} \bvec{\sigma} \cdot \nhat \, dS
\end{equation}
for any domain $\Omega$. Now consider $\Omega$ to be a section of the channel $-x_0 < x < x_0$ and spanning the entire width $-w < y < w$. Then the integral on the right-hand-side of \eqref{divthm} has two contributions: first, the total drag arising from all immersed bodies as well as from the top and bottom walls, and second, the difference between the upstream and downstream pressures (we have implicitly assumed that $x_0$ is sufficiently distant from the bodies so that $\grad \bvec{u}$ is nearly zero). Since the total integral above vanishes, we have the approximate relationship
\begin{equation}
F_D \approx 2 w \left( \pup - \pdn \right)
\end{equation}
where $F_D$ is the total drag force from all bodies and from the top and bottom walls. The error in this approximation is due to neglecting the contribution of $\grad \bvec{u}$ in the stress tensor at the upstream and downstream positions. We thus have for the permeability
\begin{equation}
k_{tot} = \frac{2 x_0 \, \qavg}{\pup-\pdn} \approx 4 x_0 w \qavg F_D^{-1}
\end{equation}
and the resistivity
\begin{equation}
r_{tot} = \frac{\pup-\pdn}{2 x_0 \, \qavg} \approx \frac{F_D}{4 x_0 w \qavg} = \frac{3 F_D}{8 x_0 w U}
\end{equation}
Where we have used $\qavg = \frac{2}{3} U$. These are the permeability and resistivity for the entire region $-x_0 < x < x_0$. In particular, the resistivity of the porous matrix will be averaged with the much lower resistivity of the buffer region. To remove the contribution from the buffer region, and thus isolate the resistivity due to the porous matrix of bodies, we can simply set $x_0 = 1$ above. With also setting $w=1$, we get
\begin{equation}
\label{resdrag1}
r_{tot} = \frac{3}{8} \frac{F_D}{U}
\end{equation}
With this simplification, I see a match between measuring resistivity via the drag \eqref{resdrag1} versus measuring resistivity directly via \eqref{resmatrix}.

% New May 5, 2022
\subsubsection{Modification for free-slip BCs}
If instead, we would like to deal with free-slip top and bottom boundaries, then $\qavg = U$, in which case
\begin{equation}
r_{tot} = \frac{1}{4 x_0} \frac{F_D}{U}
\end{equation}
where we have set $w=1$. In the post-process code, I set the default of $x_0 = 2$, which gives
\begin{equation}
r_{tot} = \frac{1}{8} \frac{F_D}{U}
\end{equation}
and this agrees pretty closely with what I see in the data.

Alternatively, if I am looking at only the matrix resistivity $r_m = x_0 r_t$, then
\begin{equation}
r_{m} = \frac{1}{4} \frac{F_D}{U}
\end{equation}
